### 2021-04-05 papers/SYRCoSE/future.tex

\section{Future Work}

There are several directions for future work.

First, we plan to apply our library to a wider range of problems. 
We are going to develop a mechanized semantics of some classical 
languages used to model concurrency, in particular 
the calculus of communicating systems~(CCS)~\cite{Milner:80} and 
$\pi$-calculus~\cite{Milner:99}.      
We also plan to continue our work on 
expressing various relaxed models of shared memory 
in terms of event structures.

Second, we want to cover other classes of event structures in our library, 
in particular bundle~\cite{Langerak:91}, flow~\cite{Boudol-Castellani:1991}
and stable~\cite{Winskel:82, Winskel:86} event structures.
We plan to use them to develop mechanized \emph{denotational} semantics 
of concurrent languages and relaxed shared memory models.

Finally, we plan to mechanize in Coq classical results 
that connect various classes of event structures%  
~\cite{Nielsen-al:1981, Boudol-Castellani:1991}. 
It would allow us to easily establish the connection between operational and denotational 
semantics of concurrent languages.