### 2021-04-05 papers/SYRCoSE/fig/regmachine-eventstruct-prog.tex

\begin{center}
    \begin{tikzpicture}
          \node (wxa) %[left of = mid, node distance = 1.1cm] 
            % at (0,1)
            {$1: \ x:=1$};

          \node (wxb) [right of = wxa, node distance = 1.9cm] 
            % at (1,1)
            {$2: \ r:=x$};

          \node (wxc) [right of = wxb, node distance = 1.9cm] 
            % at (2,1)
            {$3: \ x:=2$};

          \draw ($(wxa.north east)$) -- ($(wxa.south east)$);
          \draw ($(wxb.north west)$) -- ($(wxb.south west)$);
          \draw ($(wxb.north east)$) -- ($(wxb.south east)$);
          \draw ($(wxc.north west)$) -- ($(wxc.south west)$);
    \end{tikzpicture}
    \end{center}