### 2021-04-06 papers/SYRCoSE/future.tex
SYRCoSE: various minor fixes

\section{Future Work}

There are several directions for future work.

First, we plan to apply our library to a wider range of problems. 
We are going to develop a mechanized semantics of some classical 
languages used to model concurrency, in particular 
the calculus of communicating systems~(CCS)~\cite{Milner:80} and 
$\pi$-calculus~\cite{Milner:99}.      
We also plan to continue our work on 
expressing various relaxed models 
of shared memory~\cite{Sewell-al:CACM10, Lahav-al:POPL16, Lahav-al:PLDI17, Podkopaev-al:POPL19}
in terms of event structures.

Second, we want to cover other classes of event structures in our library, 
in particular bundle~\cite{Langerak:91}, flow~\cite{Boudol-Castellani:1991}
and stable~\cite{Winskel:82, Winskel:86} event structures.
We plan to use them to develop mechanized \emph{denotational} semantics 
of concurrent languages and relaxed shared memory models~\cite{Dodds-al:ESOP18}.

Finally, we plan to mechanize in Coq classical results 
that connect various classes of event structures%  
~\cite{Nielsen-al:1981, Boudol-Castellani:1991}. 
It would allow us to easily establish the connection between operational and denotational 
semantics of concurrent languages.