### 2021-04-06 papers/SYRCoSE/regmachine.tex

\subsection{Register Machine}

For our case study we use a simple idealized model 
of a register machine, which consists of a finite 
sequence of instructions, an instruction pointer 
and an infinite set of registers. 
The syntax of the machine's language is 
shown on~\cref{fig:regmachine-syntax}.

\begin{figure}[h!]
\[
\begin{array}{rclclll}
  P &\in& \Prog   &::=& i_1 \seq \dots \seq i_n           & \text{program}                 \\

  I &\in& \Inst     &::=&                                 & \text{instruction}             \\
              & &   & | & \assignInst{r}{v}               & \text{assign to register}      \\
              & &   & | & \opInst{\otimes}{r_1}{r_2}{r_3}  & \text{apply binary operation}  \\
              & &   & | & \jumpInst{r}{\ip{i}}            & \text{conditional jump}        \\
              & &   & | & \exitInst                       & \text{exit}                    \\
              & &   & | & \readInst{r}{x}                 & \text{read from memory}        \\
              & &   & | & \writeInst{x}{v}                & \text{write to memory}         \\

  r      &\in& \Reg     &   &                    & \text{thread-local register}   \\ 
  x      &\in& \Loc     &   &                    & \text{shared memory location}  \\ 
  v      &\in& \Z       &   &                    & \text{value}                   \\
  \otimes&\in& \BinOp   &   &                    & \text{binary operation}        \\
  \ip{i} &\in& \N       &   &                    & \text{instruction label}       \\

\end{array}
\] 
\caption{Syntax of the register machine}
\label{fig:regmachine-syntax}
\end{figure}

We first present the semantics of a single thread program.
In this semantics memory access instructions 
does not actually operate on shared memory,
but rather produce a \emph{label}
denoting the side-effect of the operation 
(see \cref{fig:label-syntax}).
This encoding allows us to decouple 
the semantics of the register machine 
from a \emph{memory model}.

\begin{figure}[h!]
\[
\begin{array}{rclclll}

  \ell &\in& \Lab  &::=&                                & \text{}             \\
              & &  & | & \rlab{x}{v}                    & \text{read of value $v$ from location $x$}   \\
              & &  & | & \wlab{x}{v}                    & \text{write of value $v$ to location $x$}    \\

\end{array}
\] 
\caption{Syntax of Labels}
\label{fig:label-syntax}
\end{figure}

The semantics is given in a form of 
\emph{labeled transition system}~%
\footnote{As we have mentioned, in our \coq development we actually use 
the monadic encoding of the operational semantics. 
The labeled transition system can be deduced from this encoding.}
State of the system consists of instruction pointer
and a map from registers to their values, 
as shown in \cref{fig:regmachine-thrdstate}. 
The rules of the semantics are standard 
(see \cref{fig:regmachine-thrdsem}).

\begin{figure}[h!]
\[
\begin{array}{rclclll}

  s          &\in& \ThrdState    &::=& \tup{\ip{i}, \regmap} &                         \\
  \ip{i}     &\in& \N            &   &                    & \text{instruction pointer} \\
  \regmap    &\in& \Reg \fun \Z  &   &                    & \text{registers mapping}   \\

\end{array}
\] 
\caption{Thread state of register machine}
\label{fig:regmachine-thrdstate}
\end{figure}

\input{fig/regmachine-thrdsem.tex}