### 2021-04-08 papers/SYRCoSE/related.tex
Related work: more grammar

\section{Related Work}

\todo{Fix Winskel1986 vs Winskel 1982 (introduction vs related work)}

Event structures were introduced by Winskel to study the semantics of 
the calculus of communicating systems~\cite{Winskel:82}. 
Several modifications of event structures~\cite{Langerak:91, Boudol-Castellani:1991}
were later proposed to tackle similar problems.

More recently event structures were applied 
in the context of relaxed memory models~%
\cite{Jeffrey-Riely:LICS16, PichonPharabod-Sewell:POPL16, Chakraborty-Vafeiadis:POPL19, Moiseenko-al:ECOOP20}.
Among this line of work, we are aware of only one paper~\cite{Moiseenko-al:ECOOP20}
that was accompanied with a mechanisation in a proof assistant. 
The authors formalised \weakestmo~\cite{Chakraborty-Vafeiadis:POPL19} 
memory memory model in \coq. However, this memory model uses 
a custom variant of event structures, that does not 
obey axioms of any conventional class of 
event structures~\cite{Winskel:82, Langerak:91, Boudol-Castellani:1991}. 
This fact makes it harder to reuse and adapt it 
to other applications of the theory.