### 2021-05-21 papers/SYRCoSE/fig/encode_prog.tex
SYRCoSE article (#74)
* feat: add SYRCoSE article

* fix: applt suggestions

* feat: add pattern of a new section

* SYRCoSE: add Makefile

* update .gitignore

* SYRCoSE: move intro into separate folder

* SYRCoSE: reworking the introduction

* SYRCoSE: finished with new intro

* SYRCoSE: minor

* SYRCoSE: add intro to event structures

* fix: .gitignore changed

* fix: 'How we encode ...' section moved to separate file

* feat: add regmachine semantics

* SYRCoSE: move intro to event structures to separate file

* feat: add part of small-step semantics section

* feat: add part of small-step

\begin{figure}[h]
  \begin{center}
    \begin{tikzpicture}

      \node (iwy) %[left of = mid, node distance = 1.1cm] 
      at (-1.3, 2.5)
      {$x:=1;$};

      \node (irx) [above = -0.5cm of iwy.west, anchor = west] 
      {$r:=x;$};

      \node (iwx) %[right of = mid, node distance = 0.8cm] 
       at (0.5, 2.5)
      {$x:=2;$};

      \node (iwz) %[right of = mid, node distance = 0.8cm] 
      at (-1.3, 1.55)
     {$y:=r;$};

     \node (iww) %[right of = mid, node distance = 0.8cm] 
      at (-1.3, 1.2)
     {$r':=y;$};

      \draw ($ (iwy.south east) + (0.1,-1.3) $) -- ($ (iwy.south east) + (0.1,0.6) $);
      \draw ($ (iwx.south west) + (-0.15,-1.3) $) -- ($ (iwx.south west) - (0.15,-0.6) $);
    \end{tikzpicture}
  \end{center}
\end{figure}