### 2021-05-21 papers/SYRCoSE/fig/intro-prog.tex

\begin{center}
\begin{tikzpicture}
      \node (wxa) %[left of = mid, node distance = 1.1cm] 
        % at (0,1)
        {$x:=1$};

      \node (wxb) [right of = wxa, node distance = 1.3cm] 
        % at (1,1)
        {$x:=2$};

      \node (wxc) [right of = wxb, node distance = 1.3cm] 
        % at (2,1)
        {$x:=3$};

      \node (rx) [below of = wxb, node distance = 0.7cm] 
        % at (1,0)
        {$r:=x$};

      \draw ($(wxa.north east)$) -- ($(wxa.south east)$);
      \draw ($(wxb.north west)$) -- ($(wxb.south west)$);
      \draw ($(wxb.north east)$) -- ($(wxb.south east)$);
      \draw ($(wxc.north west)$) -- ($(wxc.south west)$);
\end{tikzpicture}
\end{center}